Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States

### 2017-11-06 doc/manuscript/manuscript.Rmd
attempt to add discussion to intro for #15

<!-- 
Reviewer 2:
I would expand introduction a bit to describe the symptoms in greater 
details, maybe even provide some information regarding economic damage (if 
available) to nursery production. I was wondering what cultivars previous 
studies focused on – the authors provided 3 cultivars at the end of 
introduction. Maybe adding another paragraph on differences among 
cultivars – they chose 3 with different levels of resistance but how was 
that selected (assuming some preliminary data)? For someone not familiar 
with the host, I would like to have more details on both host and pathogen. 
-->
*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution [@bolton2006sclerotinia].
This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive for more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population]. 
It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia]. 
On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be yield-limiting in temperate climates [@steadman1983white].

<!-- Include discussion on symptoms and economic damage leading in to cultivar discussion? -->

Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Australia, France, and Mexico [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold [@steadman2003identification; @otto-hanson2011variation].
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.

@@

MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
Incompatibility of different MCGs resulted in formation of a barrage line accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation]. 
No MCGs were compatible with any other MCG.

>>>>

<!-- 
Reviewer 2:
I would expand introduction a bit to describe the symptoms in greater 
details, maybe even provide some information regarding economic damage (if 
available) to nursery production. I was wondering what cultivars previous 
studies focused on – the authors provided 3 cultivars at the end of 
introduction. Maybe adding another paragraph on differences among 
cultivars – they chose 3 with different levels of resistance but how was 
that selected (assuming some preliminary data)? For someone not familiar 
with the host, I would like to have more details on both host and pathogen. 
-->
*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution [@bolton2006sclerotinia].
This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive for more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population]. 
It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia].

On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be yield-limiting in temperate climates [@steadman1983white]. 
All above-ground tissues (flowers, stems, leaves, pods) are susceptible to infection, first appearing as wet lesions with white mycelial tufts, and then bleaching as the tissue senesces [@steadman1983white; @bolton2006sclerotinia].
For many years, white mold has been the most serious dry bean disease in the Northwestern United States [@otto-hanson2011variation; @knodel2012dry; @knodel2015dry; @knodel2016dry].
The impact of white mold on the dry bean industry in the Northwestern United States alone has been estimated at a loss of 140 kg/ha with just 10% disease incidence [@ramasubramaniam2008estimates].

Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Australia, France, and Mexico [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold [@steadman2003identification; @otto-hanson2011variation].
These protocols included three established cultivars used for comparison in the trials: Beryl (great northern bean, susceptible), Bunsi (aka Ex Rico, navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@tu1982tolerance; @steadman2005identification; @otto-hanson2011variation].
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.

@@

MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig. \@ref(mcg-fig)) [@cubeta2001mycelial]. 
Incompatibility of different MCGs resulted in formation of a barrage line accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation]. 
No MCGs were compatible with any other MCG.
